---
title: "p8105_hw2_hw2850"
author: "Haoyang Wang"
date: "2022-09-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

#Problem 1
```{r}
library(tidyverse)
library(dplyr)
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  dplyr::select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

```{r}
trans_ent %>% 
  dplyr::select(station_name, line) %>% 
  distinct
```

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  dplyr::select(station_name, line) %>% 
  distinct
```

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  dplyr::select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  dplyr::select(station_name, line) %>% 
  distinct
```


#Problem2
```{r} 
library(readxl)
mr_trash_wheel=read_excel("data/Trash Wheel Collection Data.xlsx", "Mr. Trash Wheel", range = "A2:N550") %>% 
  janitor::clean_names(dat = .) %>% 
  drop_na(data = ., dumpster) %>% 
  mutate(sports_balls=as.integer(sports_balls),
         dumpster = as.character(dumpster),
         name = "Mr.Trash Wheel") %>% 
  dplyr::select(.data = ., name, everything())
```

```{r}
professor_trash_wheel=read_excel("data/Trash Wheel Collection Data.xlsx", "Professor Trash Wheel") %>% 
  janitor::clean_names(dat = .) %>% 
  drop_na(data=., dumpster) %>% 
  mutate(year = as.character(year),
         dumpster = as.character(dumpster),
         name = "Professor Trash Wheel") %>% 
  dplyr::select(.data = ., name, everything())
```

```{r}
trash_wheel=bind_rows(mr_trash_wheel, professor_trash_wheel)
summary(trash_wheel)
```

```{r}
mr_trash_wheel_2020=filter(mr_trash_wheel, year=="2020")
```
* The combined data set size is `r dim(trash_wheel)`.
* The key variables are `r ls(trash_wheel)`.
* The total weight of trash collected by Professor Trash Wheel is `r sum(professor_trash_wheel$weight_tons)`.
* The total number of sports balls collected by Mr.Trash Wheel in 2020 is `r sum(mr_trash_wheel_2020$sports_balls)`.


#Problem3
```{r}
pols=read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names(dat = .) %>% 
  separate(mon, into = c("year","month","day")) %>% 
  mutate(.data = .,
    month = replace(month, month == "01", "jan"),
    month = replace(month, month == "02", "feb"),
    month = replace(month, month == "03", "mar"),
    month = replace(month, month == "04", "apr"),
    month = replace(month, month == "05", "may"),
    month = replace(month, month == "06", "jun"),
    month = replace(month, month == "07", "jul"),
    month = replace(month, month == "08", "aug"),
    month = replace(month, month == "09", "sep"),
    month = replace(month, month == "10", "oct"),
    month = replace(month, month == "11", "nov"),
    month = replace(month, month == "12", "dec"),
    president = if_else(prez_gop<1, "dem", "gop")) %>% 
  dplyr::select(.data = ., -c(day, prez_gop, prez_dem))
```
* The pols dataset has a size of `r dim(pols)`.
* The pols dataset has variables `r ls(pols)`.
* The pols dataset has a range of year `r range(pols$year)`

```{r}
snp=read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names(dat = .) %>% 
  separate(date, into = c("month","day","year")) %>%
  mutate(.data = .,
    month = replace(month, month == "1", "jan"),
    month = replace(month, month == "2", "feb"),
    month = replace(month, month == "3", "mar"),
    month = replace(month, month == "4", "apr"),
    month = replace(month, month == "5", "may"),
    month = replace(month, month == "6", "jun"),
    month = replace(month, month == "7", "jul"),
    month = replace(month, month == "8", "aug"),
    month = replace(month, month == "9", "sep"),
    month = replace(month, month == "10", "oct"),
    month = replace(month, month == "11", "nov"),
    month = replace(month, month == "12", "dec")) %>% 
  mutate(.data = ., year = as.numeric(year),
         y = if_else(year < 22, 2000, 1900),
         year = year + y,
         year = as.character(year),
         ) %>% 
  dplyr::select(.data = .,-day,-y) %>% 
  relocate(year, month) 
```
* The snp dataset has a size of `r dim(snp)`.
* The snp dataset has variables `r ls(snp)`.
* The snp dataset has a range of year `r range(snp$year)`

```{r}
unemploy=read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv", col_types = "ccccccccccccc") %>% 
  janitor::clean_names(dat = .) 

unemploy_tidy_data=
  pivot_longer(
    unemploy,
    jan:dec,
    names_to = "month",
    values_to = "unemployment_value") %>% 
  mutate(.data = ., unemployment_value = as.numeric(unemployment_value))
```
* The unemployment dataset has a size of `r dim(unemploy_tidy_data)`.
* The unemployment dataset has variables `r ls(unemploy_tidy_data)`.
* The unemployment dataset has a range of year `r range(unemploy_tidy_data$year)`.

```{r}
snp_to_pols=
  left_join(pols, snp, by = c("year", "month"))

unemployment_snp_pols=
  left_join(snp_to_pols, unemploy_tidy_data, by = c("year", "month"))
```
* The final dataset unemployment_snp_pols has a dimension `r dim(unemployment_snp_pols)`.
* The final dataset has variables `r ls(unemployment_snp_pols)`.
* The final dataset has a range of year `r range(unemployment_snp_pols$year)`.
